---
title: "Censo 2022"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## Endereço das Planilhas

[Dicionario de Dados Agregados por Setor Censitário](https://ftp.ibge.gov.br/Censos/Censo_Demografico_2022/Agregados_por_Setores_Censitarios/dicionario_de_dados_agregados_por_setores_censitarios.xlsx)

[Censo 2022 - Dados de Alfabetização por Municipio](https://www.ibge.gov.br/estatisticas/downloads-estatisticas.html?caminho=Censos/Censo_Demografico_2022/Agregados_por_Setores_Censitarios/Agregados_por_Municipio_csv/)

[Censo 2010 - Dados de Alfabetização por Municipio]()


[Publicação do IBGE Com Resultados Relacionados a Alfabetização](https://biblioteca.ibge.gov.br/visualizacao/periodicos/3108/cd_2022_alfabetizacao.pdf)

É apresentada aqui uma breve análise dos resultados do quesito sobre alfabetização 
presente no Questionário Básico do Censo Demográfico de 2022. Para tanto, recorre-se 
ao indicador de taxa de alfabetização das pessoas de 15 anos ou mais de idade e 
ao seu **complemento**, taxa de analfabetismo das pessoas de 15 anos ou mais de idade, 
conforme sugerido pelo documento “Princípios e Recomendações para os Censos Populacionais”, 
elaborado pela Divisão de Estatística da ONU. Os resultados são analisados em 
seis seções: desagregações por grupos de idade, cor ou raça, sexo, tamanho dos 
municípios, regiões geográficas e pessoas indígenas.

Os resultados do Censo Demográfico 2022 mostram que, no País, havia 163 milhões
de pessoas de 15 anos ou mais de idade, das quais 151,5 milhões sabiam ler e escrever
um bilhete simples e 11,4 milhões não sabiam sequer ler e escrever um bilhete simples.
A partir desses totais populacionais, a taxa de alfabetização foi 93,0% em 2022 e, 
consequentemente, a taxa de analfabetismo foi 7,0% deste contingente populacional. 

*Leitura dos Dados*

```{r Leitura}
library("tidyverse")
library("data.table")
library("abjutils")
#setwd("~/Github/DadosCenso2022")
url <- "~/GitHub/Censo2022/Dados/Agregados_por_municipios_alfabetizacao_BR.csv"
#url <- "~/GitHub/Censo2022/Dados/Agregados_por_setores_alfabetizacao_BR.csv"

dt <- fread(url,
            encoding = "Latin-1",
            sep = ";",
            header = TRUE)

```


###  





Write the R code required to add two plus two:

```{r two-plus-two, exercise=TRUE}

```

### Exercise with Code

*Here's an exercise with some prepopulated code as well as `exercise.lines = 5` to provide a bit more initial room to work.*

Now write a function that adds any two numbers and then call it:

```{r add-function, exercise=TRUE, exercise.lines = 5}
add <- function() {
  
}
```

## Topic 2

### Exercise with Hint

*Here's an exercise where the chunk is pre-evaluated via the `exercise.eval` option (so the user can see the default output we'd like them to customize). We also add a "hint" to the correct solution via the chunk immediate below labeled `print-limit-hint`.*

Modify the following code to limit the number of rows printed to 5:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-hint}
head(mtcars)
```

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question("Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```

